Simulation of Congenital Heart Defect Corrective Surgeries Using Thin Shell Elements

### 2012-03-01 Doc/Miccai/introduction.tex
Responded to Christian's comments.
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@359 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

To improve planning of such pediatric cardiac surgeries it is possible to create 3-dimensional heart models out of acquired image data REF and to use them as input for surgical simulation systems. However, most present simulation systems focus on low-level surgical procedures, i.e. performing actual incisions and suturing with special input devices in real-time REF. Their purpose is either by means of educational training or planning access strategies to reach the malformations of the heart or great arteries. In such {\itshape reactive} \christian{interactive ?} simulation systems it is more important to enable simulations in real-time than to simulate tissue and blood vessels in a physically accurate manner.
\christian{We could cite some existing paper about interactive simulation for training}

To be able to predict the result of complex surgical procedures, a physically correct simulation is mandatory. In \cite{Li2009} is presented a {\itshape predictive} surgical simulation system for bidirectional Glenn shunts and basically similar surgeries. They use a hybrid model to simulate deformations of blood vessels by binding a surface mesh to a reference Cosserat rod REF elastically. While the mesh models surface details of blood vessels, the Cosserat rod models global deformations, i.e. stretching, bending, twisting and shearing. Joining of blood vessels end-to-end or end-to-side is realized by attaching contracting springs. This approach adds surplus energy to the physical system and special care has to be taken during the simulation to compensate for that. The software performs low-level surgical procedures automatically with only a minimum of interaction required which allows the surgeon to focus on the actual overall surgical procedure. Given the input of the surgeon, a complete simulation takes only seconds to produce a predictive result for further examination and evaluation of possibly undesired collisions or deformations of blood vessels. 
Although the results are promising, the method is limited to operations on uniform tubular structures that do not arbitrarily change the surface. Surgical procedures heavily changing surface details of blood vessels, i.e. patches for correction of complicated coarctations and stenosis, are not possible with this hybrid model since the physical behavior of blood vessels is simulated by the Cosserat rod and its {\itshape binding} to the surface mesh.
\christian{I think this paragraph is much too long and I think that our paper should not be presented in opposition to \cite{Li2009} but in the continuity of this work !!  I put a correction of this in a commit 2 days ago, could we try to merge ?}

In this paper we present an alternative, improved approach using a simulation model solely based on elastic thin shell elements. Our method is not limited to well-formed blood vessels and captures all low-level procedures necessary for aortoplasty. By using a relaxation process we can arbitrarily join or attach blood vessels and patches without using springs that would add surplus energy to the physical system. We prove that simulation time is still reasonable or even decreased by taking special care of the overall tubular structure of blood vessels during generation of thin shell elements.

>>>>

To improve planning of such pediatric cardiac surgeries it is possible to create 3-dimensional heart models out of acquired image data REF and to use them as input for surgical simulation systems. However, most present simulation systems focus on low-level surgical procedures, i.e. performing actual incisions and suturing with special input devices in real-time REF. Their purpose is either by means of educational training or planning access strategies to reach the malformations of the heart or great arteries. In such {\itshape reactive} \christian{interactive ?} \stefan{reactive is the term used by current papers} simulation systems it is more important to enable simulations in real-time than to simulate tissue and blood vessels in a physically accurate manner.
\christian{We could cite some existing paper about interactive simulation for training} \stefan{We can cite two papers here I will add today. They will be put to the second REF in this paragraph.}

To be able to predict the result of complex surgical procedures, a physically correct simulation is mandatory. In \cite{Li2009} is presented a {\itshape predictive} surgical simulation system for bidirectional Glenn shunts and basically similar surgeries. They use a hybrid model to simulate deformations of blood vessels by binding a surface mesh to a reference Cosserat rod REF elastically. While the mesh models surface details of blood vessels, the Cosserat rod models global deformations, i.e. stretching, bending, twisting and shearing. Joining of blood vessels end-to-end or end-to-side is realized by attaching contracting springs. This approach adds surplus energy to the physical system and special care has to be taken during the simulation to compensate for that. The software performs low-level surgical procedures automatically with only a minimum of interaction required which allows the surgeon to focus on the actual overall surgical procedure. Given the input of the surgeon, a complete simulation takes only seconds to produce a predictive result for further examination and evaluation of possibly undesired collisions or deformations of blood vessels. 
Although the results are promising, the method is limited to operations on uniform tubular structures that do not arbitrarily change the surface. Surgical procedures heavily changing surface details of blood vessels, i.e. patches for correction of complicated coarctations and stenosis, are not possible with this hybrid model since the physical behavior of blood vessels is simulated by the Cosserat rod and its {\itshape binding} to the surface mesh.
\christian{I think this paragraph is much too long and I think that our paper should not be presented in opposition to \cite{Li2009} but in the continuity of this work !!  I put a correction of this in a commit 2 days ago, could we try to merge ?} \stefan{I already took your comments into account and tried to present our approach as a consequent continuity of this work. The paragraph was shortened by some sentences also (see the comments in the bottom of the Tex-File. I'm not sure if we should further shorten this paragraph \emph{especially} when we present our work as continuity. Because that paragraph is also an introduction for our work then, isn't it? In the end the paragraph says what can be improved to justify our improved approach. Maybe when we just delete the "alternative" in the next paragpraph?}

In this paper we present an \stefan{(}alternative\stefan{)}, improved approach using a simulation model solely based on elastic thin shell elements. Our method is not limited to well-formed blood vessels and captures all low-level procedures necessary for aortoplasty. By using a relaxation process we can arbitrarily join or attach blood vessels and patches without using springs that would add surplus energy to the physical system. We prove that simulation time is still reasonable or even decreased by taking special care of the overall tubular structure of blood vessels during generation of thin shell elements.

### 2012-03-01 Doc/Miccai/shell-model.tex
comments on intro and section2
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@358 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b
Responded to Christian's comments.
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@359 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\stefan{Alternative start of this section: In congenital heart defect (CHD) corrective surgeries blood vessels respond to plastic interventions with a certain amount of deformation. Surgeons try to find an optimal approach to correct defects with a minimum of deformations and collisions that would have unnecessary negative impact on blood flow or that may lead to problems during growth of the patient. To be able to predict results for different surgical approaches a physical approriate model of blood vessels is mandatory. Furthermore, computation time of simulations must be low enough to enable interactive and usable planning software. Options on the choice of model are limited. Mass-spring models allow for fast computation but do not correlate to a physical accurate representation of reality and therefore are inappropriate for predictive simulations. On the contrary, simulation models based on the formulation of finite elements, either 3-dimensional elements or thin shell elements, are commonly used when high physical accuracy is required but are well known to be computationally expensive. The hybrid blood vessel model of REF strikes a balance but is too constrained to be used as a truly universal approach for all kinds of surgical procedures used in CHD corrective surgery. We propose a model based on thin shell elements that still allows for fast computation with a reasonable number of elements.}

>>>>

\stefan{Alternative start of this section: In congenital heart defect (CHD) corrective surgeries blood vessels respond to plastic interventions with a certain amount of deformation. Surgeons try to find an optimal approach to correct defects with a minimum of deformations and collisions that would have unnecessary negative impact on blood flow or that may lead to problems during growth of the patient. To be able to predict results for different surgical approaches a physical approriate model of blood vessels is mandatory. Furthermore, computation time of simulations must be low enough to enable interactive and usable planning software. Options on the choice of model are limited. Mass-spring models allow for fast computation but do not correlate to a physical accurate representation of reality and therefore are inappropriate for predictive simulations. On the contrary, simulation models based on the formulation of finite elements, either 3-dimensional elements or thin shell elements, are commonly used when high physical accuracy is required but are well known to be computationally expensive. The hybrid blood vessel model of REF strikes a balance but is too constrained to be used as a truly universal approach for all kinds of surgical procedures used in CHD corrective surgery. We propose a model based on thin shell elements that still allows for fast computation with a reasonable number of elements.}
\christian{I think we really need to do a merge because there are interesting things in both section but in both cases, we are missing references !!
One other important feature: I think it is important to put some illustrations of the surgery either here or in the intro.} \stefan{Yes, we truly have to merge here and add references. My addition focused more on English language.}
